---
title: "Melanoma_SubPOP"
output: html_notebook
---

Loading data
```{r Load Data}

source("/Users/andrewleduc/Desktop/nPOP_Paper/code/functions_parameters.R")

#t7_Monocytes <- read.csv('../dat/Proccessed_data/t7_Monocytes.csv',row.names = 1)
#t7_Melanomas <- read.csv('../dat/Proccessed_data/t7_Melanomas.csv',row.names = 1)

# Databases for Protein Set Enrichment Analysis
GO_db <- read.delim("../dat/raw_data/GOA_db.txt", sep = " ")

# Data base for converting Gene Names to Proteins
convert <- read.delim('../dat/raw_data/human_uniprot_db.txt')
convert$Gene.names <- as.character(convert$Gene.names)
convert$Gene.names2 <- gsub(' .*','',convert$Gene.names)

#load Complexes database
comps <- read.delim('../dat/raw_data/coreComplexes.txt')
comps <- comps %>% distinct(ComplexName, .keep_all = T)

#Load Meta data for updating
meta <- read.csv('../dat/Proccessed_data/meta.csv',row.names = 1)
meta_mel <- meta %>% filter(celltype == 'm')
meta_mon <- meta %>% filter(celltype == 'u')
meta_mel_l <- meta %>% filter(sub =='C1')
meta_mel_s <- meta %>% filter(sub =='C2')

t7 <- read.csv('../dat/Proccessed_data/t7.csv',row.names = 1)
t7_Monocytes <- t7 %>% dplyr::select(intersect(colnames(t7),meta_mon$id))
t7_Melanomas <- t7 %>% dplyr::select(intersect(colnames(t7),meta_mel$id))
t7_Melanomas_small <-t7 %>% dplyr::select(intersect(colnames(t7),meta_mel_s$id))
t7_Melanomas_large <-t7 %>% dplyr::select(intersect(colnames(t7),meta_mel_l$id))


```

Here we are Computing PCA in the space of the Melanoma cells where we have noticed 2 populations. We overlay markers from Emert et al. (2021) paper on the same melanoma cells that are predictive of either drug resistance or drug suceptibility. We compute p values and notice strong allignment with our two observed clusters

```{r  Sub population Analysis of Melanoma Cells Drug Resistance}
#Load markers for Drug resistance from Emert et al. (2021)


raj <- read.csv('../dat/raw_data/Melanoma_diff_exp_raj.csv')
raj <- raj %>% left_join(convert,by = c('gene_name' = 'Gene.names2'))
raj <- raj %>% dplyr::filter(padj < .1)

Drug_resist <- raj %>% dplyr::filter(log2FoldChange >1.5)
Main_pop <- raj %>% dplyr::filter(log2FoldChange < 0)


Drug_resist <- Drug_resist %>% filter(Entry %in% rownames(t7))
#PCA

mat.sc.imp <- as.matrix(t7_Melanomas)
mat.sc.imp <- cr_norm_log(mat.sc.imp)

X.m <-  mat.sc.imp
pca.imp.cor <- cor(X.m, use = 'pairwise.complete.obs',method = c('pearson'))

# PCA
sc.pca<-eigen(pca.imp.cor)
scx<-as.data.frame(sc.pca$vectors)
colnames(scx)<-paste0("PC",1:ncol(scx))
scx$cells<-colnames(pca.imp.cor)

# Percent of variance explained by each principle component
pca_var <- sc.pca$values
percent_var<- pca_var/sum(pca_var)*100
plot(1:length(percent_var), percent_var, xlab="PC", ylab="% of variance explained")

# Map meta data
pca.melt <- melt(scx); colnames(pca.melt)<-c("id","pc","value")
ev.melt.uniqueID_lim<-meta[,c(4:6)]

pca.melt <- left_join(pca.melt,ev.melt.uniqueID_lim,by= 'id')
pca.display <- reshape2::dcast(pca.melt, id + celltype + channel ~ pc, value.var = "value", fill=NA)

PCx<-"PC1"
PCy<-"PC2"


# Color PCA by markers of Resistance

#t7_pif_filt_mel <-as.data.frame(t7_pif_filt) %>%dplyr::select(colnames(t7_Melanomas))
Res  <- t7_Melanomas %>% dplyr::filter(rownames(t7_Melanomas) %in% Drug_resist$Entry)
Main <- t7_Melanomas %>% dplyr::filter(rownames(t7_Melanomas) %in% Main_pop$Entry)


pca.display$Resistance_Markers <- colMedians(as.matrix(Res),na.rm = T)
pca.display$Resistance_Markers[pca.display$Resistance_Markers > 1] <- 1
pca.display$Resistance_Markers[pca.display$Resistance_Markers < -1] <- -1
pca.display$Main_Population_Markers <- colMedians(as.matrix(Main),na.rm = T)
pca.display$Main_Population_Markers[pca.display$Main_Population_Markers > 1] <- 1
pca.display$Main_Population_Markers[pca.display$Main_Population_Markers < -1] <- -1

sp4 <- ggscatter(pca.display, color = 'Resistance_Markers' ,x =PCx, y = PCy, size = 5, alpha=0.5) +
  xlab("") +
  ylab("") +
  font("ylab",size=0) +
  font("xlab",size=0) +
  font("xy.text", size=0)+
  rremove("legend") + 
  ggtitle('Resistance Markers') + 
  scale_color_gradient2(midpoint = 0, low = "blue", mid = "white",
                          high = "red",name = '') + theme(plot.title = element_text(size = 20))

bp_m <-ggscatter(pca.display, color = 'Main_Population_Markers' ,x =PCx, y = PCy, size = 8, alpha=0.5) +
  xlab(paste0(PCx,"  (", round(percent_var[1],0),"%)")) +
  ylab(paste0(PCy,"  (", round(percent_var[2],0),"%)")) +
  font("ylab",size=30) +
  font("xlab",size=30) +
  font("xy.text", size=20)+
  #rremove("legend") +
  ggtitle('Main Population Markers') + 
  theme(legend.key.height =  unit(.35, "cm"),legend.key.width  =  unit(3, "cm")) + 
  scale_color_gradient2(midpoint = 0, low = "blue", mid = "white",
                          high = "red",name = '') + theme(plot.title = element_text(size = 25))




#Get Identifiers for subpopulations based off the GSEA
Melanoma_Small <- pca.display %>% dplyr::filter(PC1 < -.07)
Melanoma_Large <- pca.display %>% dplyr::filter(PC1 > -.07)

t7_Melanomas_Small <- t7_Melanomas %>% dplyr::select(Melanoma_Small$id)
t7_Melanomas_Large <- t7_Melanomas %>% dplyr::select(Melanoma_Large$id)

#meta$sub <- NA
#meta$sub[meta$id %in% colnames(t7_Melanomas_Large)] <- 'C1'
#meta$sub[meta$id %in% colnames(t7_Melanomas_Small)] <- 'C2'

#write.csv(t7_Melanomas_Small,'../dat/Proccessed_data/t7_Melanomas_Small.csv',row.names = T)
#write.csv(t7_Melanomas_Large,'../dat/Proccessed_data/t7_Melanomas_Large.csv',row.names = T)
#write.csv(meta,'../dat/Proccessed_data/meta.csv',row.names = T)

#calc p val for marker clusters
Res_small  <- t7_Melanomas_Small %>% filter(rownames(t7_Melanomas_Small) %in% Drug_resist$Entry)
Res_large  <- t7_Melanomas_Large %>% filter(rownames(t7_Melanomas_Large) %in% Drug_resist$Entry)
Res_small <- melt(Res_small)
Res_small$variable <- 'Small'
Res_large<- melt(Res_large)
Res_large$variable <- 'Large'
res_p <- rbind(Res_large,Res_small)
res_p <- res_p %>% filter(is.na(value)==F)
kruskal.test(value ~ variable, data = res_p)
#3.864e-05

Main_small <- t7_Melanomas_Small %>% filter(rownames(t7_Melanomas_Small) %in% Main_pop$Entry)
Main_large <- t7_Melanomas_Large %>% filter(rownames(t7_Melanomas_Large) %in% Main_pop$Entry)
Main_small<- melt(Main_small)
Main_small$variable <- 'Small'
Main_large <- melt(Main_large)
Main_large$variable <- 'Large'
main_p <- rbind(Main_large,Main_small)
main_p <- main_p %>% filter(is.na(value)==F)
kruskal.test(value ~ variable, data = main_p)
kruskal.test(value ~ variable, data = res_p)
bp_m
#2.2e-16



```
Here we analyze cluster further and and compute protein set enrichment analysis between the two cluster. We identify many significant gene sets related to cell division.
```{r  Sub population Analysis of Melanoma Cells GSEA}
## GSEA on Sub population

Melanoma_GSEA <- t7_Melanomas
Melanoma_GSEA$prot <- rownames(Melanoma_GSEA)
Melanoma_GSEA <- melt(Melanoma_GSEA,id.vars = 'prot')
Melanoma_GSEA$Condition <- NA
Melanoma_GSEA$Condition[Melanoma_GSEA$variable %in% Melanoma_Small$id] <- 'Small'
Melanoma_GSEA$Condition[Melanoma_GSEA$variable %in% Melanoma_Large$id] <- 'Large'
Melanoma_GSEA <- Melanoma_GSEA %>% filter(is.na(value)==F)
#Melanoma_GSEA$variable <- NULL

colnames(Melanoma_GSEA)<- c('Protein','SampID','Intensity','Condition')


GSEA_out <- Population_GSEA(Melanoma_GSEA,GO_db,2)
GSEA_out <- GSEA_out %>% filter(is.na(pVal)==F)

# Multiple hypothesis testing correction
GSEA_out$qVal <- p.adjust(GSEA_out$pVal, method = "BH")
GSEA_out <- GSEA_out %>% filter(qVal < .0001)
GSEA_out <- GSEA_out %>% filter(numberOfMatches > 2)
GSEA_out$dif <- GSEA_out$Cond1med_int - GSEA_out$Cond2med_int


#write.csv(GSEA_out,'../dat/Output_tables/Melanoma_SubPop_PSEA.csv')

GO_db_set1 <- GO_db %>% filter(GO_term_name == 'protein secretion' )
Gene_set1 <- t7_Melanomas %>% filter(rownames(t7_Melanomas) %in% GO_db_set1$Uniprot)
pca.display$p53 <- colMedians(as.matrix(Gene_set1),na.rm = T)
pca.display$p53[pca.display$p53 > 1] <- 1
pca.display$p53[pca.display$p53 < -1] <- -1
sp1 <-ggscatter(pca.display, color = 'p53' ,x =PCx, y = PCy, size = 5, alpha=0.5) +
  xlab("") +
  ylab("") +
  font("ylab",size=0) +
  font("xlab",size=0) +
  font("xy.text", size=0)+
  rremove("legend") + 
  ggtitle('protein secretion') + theme(plot.title = element_text(size = 14)) +
  scale_color_gradient2(midpoint = .2, low = "blue", mid = "white",
                          high = "red",name = '')+ theme(plot.title = element_text(size = 20))

GO_db_set1 <- GO_db %>% filter(GO_term_name == 'cellular senescence' )
Gene_set1 <- t7_Melanomas %>% filter(rownames(t7_Melanomas) %in% GO_db_set1$Uniprot)
pca.display$cell_sens <- colMedians(as.matrix(Gene_set1),na.rm = T)
pca.display$cell_sens[pca.display$cell_sens > 1] <- 1
pca.display$cell_sens[pca.display$cell_sens < -1] <- -1
sp2 <-ggscatter(pca.display, color = 'cell_sens' ,x =PCx, y = PCy, size = 5, alpha=0.5) +
  xlab("") +
  ylab("") +
  font("ylab",size=0) +
  font("xlab",size=0) +
  font("xy.text", size=0)+
  rremove("legend") +
  ggtitle('cellular senesence') +
  scale_color_gradient2(midpoint = .0, low = "blue", mid = "white",
                          high = "red",name = '') + theme(plot.title = element_text(size = 20))

GO_db_set1 <- GO_db %>% filter(GO_term_name == 'positive regulation of cell cycle arrest')
Gene_set1 <- t7_Melanomas %>% filter(rownames(t7_Melanomas) %in% GO_db_set1$Uniprot)
pca.display$resp_c <- colMedians(as.matrix(Gene_set1),na.rm = T)
pca.display$resp_c[pca.display$resp_c > 1] <- 1
pca.display$resp_c[pca.display$resp_c < -1] <- -1
sp3 <-ggscatter(pca.display, color = 'resp_c' ,x =PCx, y = PCy, size = 5, alpha=0.5) +
  xlab("") +
  ylab("") +
  font("ylab",size=0) +
  font("xlab",size=0) +
  font("xy.text", size=0)+
  rremove("legend") + 
  ggtitle('positive regulation of cell cycle arrest') +
  scale_color_gradient2(midpoint = .2, low = "blue", mid = "white",
                          high = "red",name = '') + theme(plot.title = element_text(size = 20))



GO_db_set1 <- GO_db %>% filter(GO_term_name == 'G2/M transition of mitotic cell cycle' )
Gene_set1 <- t7_Melanomas %>% filter(rownames(t7_Melanomas) %in% GO_db_set1$Uniprot)
pca.display$G2 <- colMedians(as.matrix(Gene_set1),na.rm = T)
pca.display$G2[pca.display$G2 > 1] <- 1
pca.display$G2[pca.display$G2 < -1] <- -1
bp1 <-ggscatter(pca.display, color = 'G2' ,x =PCx, y = PCy, size = 5, alpha=0.5) +
  xlab("") +
  ylab("") +
  font("ylab",size=0) +
  font("xlab",size=0) +
  font("xy.text", size=0)+
  rremove("legend") + 
  ggtitle('G2/M transition of mitotic cell cycle') + 
  scale_color_gradient2(midpoint = .0, low = "blue", mid = "white",
                          high = "red",name = '') + theme(plot.title = element_text(size = 20))

GO_db_set1 <- GO_db %>% filter(GO_term_name == 'microtubule-based process' )
Gene_set1 <- t7_Melanomas %>% filter(rownames(t7_Melanomas) %in% GO_db_set1$Uniprot)
pca.display$tub <- colMedians(as.matrix(Gene_set1),na.rm = T)
pca.display$tub[pca.display$tub > 1] <- 1
pca.display$tub[pca.display$tub < -1] <- -1
bp2 <-ggscatter(pca.display, color = 'tub' ,x =PCx, y = PCy, size = 5, alpha=0.5) +
  xlab("") +
  ylab("") +
  font("ylab",size=0) +
  font("xlab",size=0) +
  font("xy.text", size=0)+
  rremove("legend") +
  ggtitle('microtubule-based process') + 
  scale_color_gradient2(midpoint = .0, low = "blue", mid = "white",
                          high = "red",name = '')+ theme(plot.title = element_text(size = 20))



GO_db_set1 <- GO_db %>% filter(GO_term_name == 'regulation of cyclin-dependent protein kinase activity')
Gene_set1 <- t7_Melanomas %>% filter(rownames(t7_Melanomas) %in% GO_db_set1$Uniprot)
pca.display$gluc <- colMedians(as.matrix(Gene_set1),na.rm = T)
pca.display$gluc[pca.display$gluc > 1] <- 1
pca.display$gluc[pca.display$gluc < -1] <- -1
bp3 <-ggscatter(pca.display, color = 'gluc' ,x =PCx, y = PCy, size = 5, alpha=0.5) +
  xlab("") +
  ylab("") +
  font("ylab",size=0) +
  font("xlab",size=0) +
  font("xy.text", size=0)+
  rremove("legend") + 
  ggtitle('cyclin-dependent kinase activity') + 
  scale_color_gradient2(midpoint = .0, low = "blue", mid = "white",
                          high = "red",name = '')+ theme(plot.title = element_text(size = 20))

GO_db_set1 <- GO_db %>% filter(GO_term_name == 'ubiquitin ligase complex')
Gene_set1 <- t7_Melanomas %>% filter(rownames(t7_Melanomas) %in% GO_db_set1$Uniprot)
pca.display$ccarr <- colMedians(as.matrix(Gene_set1),na.rm = T)
pca.display$ccarr[pca.display$ccarr > 1] <- 1
pca.display$ccarr[pca.display$ccarr < -1] <- -1
bp4 <-ggscatter(pca.display, color = 'ccarr' ,x =PCx, y = PCy, size = 5, alpha=0.5) +
  xlab("") +
  ylab("") +
  font("ylab",size=0) +
  font("xlab",size=0) +
  font("xy.text", size=0)+
  rremove("legend") + 
  ggtitle('ubiquitin ligase complex') + 
  scale_color_gradient2(midpoint = .0, low = "blue", mid = "white",
                          high = "red",name = '') + theme(plot.title = element_text(size = 20))

##plot
s <- (bp_m + (sp4 | sp3  | sp2 | sp1)/(bp1 | bp2 | bp3 | bp4)) + plot_layout(ncol=2,widths = c(1,2))
s
#ggsave('/Users/andrewleduc/Desktop/nPOP_Paper/figs/SubPop/SubPOP.png',plot = s,height = unit(6, "cm"),width = unit(20, "cm"))

#ggsave('/Users/andrewleduc/Desktop/test_fig3_2.png',plot =(comp_mon | comp_mel),height = unit(5, "cm"),width = unit(20, "cm"))
```

In the next two sections, we look to suggest the presence of complexes in our single cell data by computing similarity (cosine distance normalized for # of pairwise data points in comparison of two vectors) between proteins from a complex, and compare it to the distribution of all pairwise euclidean distances.
```{r Protein Complex Analysis Monocyte}

# Compute Null Distribution for all pairwise euclidean distances in melanoma 
Monocyte_null_dist <- Distance_euclid_py(as.matrix(t7_Monocytes),10)
Monocyte_null_dist$Condition <- " All Proteins"
Monocyte_null_dist <- Monocyte_null_dist %>% dplyr::select(Condition,value)

Monocyte_Complexes <- data.frame(Complex_name = character(), pVal = numeric(),dist = numeric(),number_prots = numeric(), stringsAsFactors = FALSE)

#calculate distances for each complex in dataset
count <- 0
for(i in 1:nrow(comps)){
  #GO_term <- unique_GO_terms[i]
  #GO_db_lim <- GO_db %>% dplyr::filter(GO_term_name == GO_term)
  comp_list <- unlist(str_split(comps$subunits.UniProt.IDs.[i],';'))
  GO_mat <- t7_Monocytes %>% filter(rownames(t7_Monocytes) %in% comp_list)
  if (nrow(GO_mat) > 2){
    GO_mat <- as.matrix(GO_mat)
    #GO_weight <- Mon_weights[rownames(Mon_weights)%in%rownames(GO_mat),]
    GO_cor <- Distance_euclid_py(GO_mat,10)
    if(nrow(GO_cor) > 3){
      count <- count+1
      GO_cor$Condition <- 'Exp'
      GO_cor <- GO_cor %>% dplyr::select(Condition,value)
      Null_plus_dat <- rbind(Monocyte_null_dist,GO_cor)
      aov_out <- aov(value ~ Condition, data = Null_plus_dat)
      aov_out<- summary(aov_out)
      aov_out <- data.frame(aov_out[[1]])
      
      Monocyte_Complexes[count,1] <- comps$ComplexName[i]
      Monocyte_Complexes[count,2] <- aov_out[1,5]
      Monocyte_Complexes[count,3] <- median(GO_cor$value)
      Monocyte_Complexes[count,4] <- nrow(GO_mat)
    }
  }
}

Monocyte_Complexes$qVal <- p.adjust(Monocyte_Complexes$pVal, method = 'BH')
Monocyte_Complexes <- Monocyte_Complexes %>% filter(qVal < .1)
Monocyte_Complexes <- Monocyte_Complexes %>% filter(dist < median(Monocyte_null_dist$value))

#write.csv(Monocyte_Complexes,'../dat/Output_tables/MonocyteComplex.csv')

# Manually select Complexes to Plot


Mon_comp_list <- c("BBS-chaperonin complex","40S ribosomal subunit, cytoplasmic","TNF-alpha/NF-kappa B signaling complex 6")

Mon_comp_plot <- Monocyte_null_dist

for (X in Mon_comp_list){
  
  comp_list <- unlist(str_split(comps$subunits.UniProt.IDs.[comps$ComplexName == X],';'))
  GO_mat <- t7_Monocytes %>% filter(rownames(t7_Monocytes) %in% comp_list)
  GO_mat <- as.matrix(GO_mat)
  GO_cor <- Distance_euclid_py(GO_mat,10)
  GO_cor$Condition <- X
  GO_cor <- GO_cor %>% dplyr::select(Condition,value)
  Mon_comp_plot <- rbind(Mon_comp_plot,GO_cor)
  
}

Mon_comp_plot$Condition[Mon_comp_plot$Condition == "40S ribosomal subunit, cytoplasmic"] <- 'z ribosomal subunit, cytoplasmic'

Mon_comp_plot$Condition[Mon_comp_plot$Condition == "TNF-alpha/NF-kappa B signaling complex 6"] <- 'TNF-alpha/NF-kappa B signaling complex'

comp_mon <- ggplot(Mon_comp_plot,aes(x = log10(value),y = Condition)) +
  geom_density_ridges(aes(fill = Condition),stat='binline' ,bins = 35,rel_min_height = 0.01,scale = 2,alpha = .0,color=NA) +
  scale_fill_manual(values = c("#808080", "#007A39","#007A39","#007A39")) +
  theme(legend.position = "none") + 
  geom_density_ridges(data=subset(Mon_comp_plot, Condition %in% Mon_comp_list[1]),
                      aes(fill = Condition),jittered_points = TRUE,rel_min_height = 0.01,scale = .15,alpha = .5)+
  theme_bw() + coord_cartesian(xlim = c(-2,0)) + 
  ggtitle('Monocyte Complexes') + ylab('') + 
  xlab('Log10(Euclidean Distance)') +theme(legend.position = "none",axis.text=element_text(size=16),axis.title = element_text(size = 16),title = element_text(size = 18))+
  geom_density_ridges(data=subset(Mon_comp_plot, Condition == 'TNF-alpha/NF-kappa B signaling complex'),
                      aes(fill = Condition),jittered_points = TRUE,rel_min_height = 0.01,scale = .5,alpha = .5)+
  geom_density_ridges(data=subset(Mon_comp_plot, Condition == 'z ribosomal subunit, cytoplasmic'),
                      aes(fill = Condition),jittered_points = TRUE,point_alpha=.1,rel_min_height = 0.01,scale = .3,alpha = .5)+
  geom_density_ridges(data=subset(Mon_comp_plot, Condition == ' All Proteins'),
                      aes(fill = Condition),jittered_points = FALSE,rel_min_height = 0.02,scale = .7,alpha = .7) + coord_cartesian(xlim = c(-2,-.4))
comp_mon

```

```{r Protein Complex Analysis Melanomas}
Melanoma_null_dist <- Distance_euclid_py(as.matrix(t7_Melanomas),10)
Melanoma_null_dist$Condition <- " All Proteins"
Melanoma_null_dist <- Melanoma_null_dist %>% dplyr::select(Condition,value)


Melanoma_Complexes <- data.frame(Complex_name = character(), pVal = numeric(),dist = numeric(),number_prots = numeric(), stringsAsFactors = FALSE)

count <- 0
for(i in 1:nrow(comps)){
  #GO_term <- unique_GO_terms[i]
  #GO_db_lim <- GO_db %>% dplyr::filter(GO_term_name == GO_term)
  comp_list <- unlist(str_split(comps$subunits.UniProt.IDs.[i],';'))
  GO_mat <- t7_Melanomas %>% filter(rownames(t7_Melanomas) %in% comp_list)
  if (nrow(GO_mat) > 2){
    GO_mat <- as.matrix(GO_mat)
    GO_cor <- Distance_euclid_py(GO_mat,10)
    if(nrow(GO_cor) > 2){
      count <- count+1
      GO_cor$Condition <- 'Exp'
      GO_cor <- GO_cor %>% dplyr::select(Condition,value)
      Null_plus_dat <- rbind(Melanoma_null_dist,GO_cor)
      aov_out <- aov(value ~ Condition, data = Null_plus_dat)
      aov_out<- summary(aov_out)
      aov_out <- data.frame(aov_out[[1]])
      
      Melanoma_Complexes[count,1] <- comps$ComplexName[i]
      Melanoma_Complexes[count,2] <- aov_out[1,5]
      Melanoma_Complexes[count,3] <- median(GO_cor$value)
      Melanoma_Complexes[count,4] <- nrow(GO_mat)
      
      

    }
  }
}

Melanoma_Complexes$qVal <- p.adjust(Melanoma_Complexes$pVal, method = 'BH')
Melanoma_Complexes <- Melanoma_Complexes %>% filter(qVal < .1)
Melanoma_Complexes <- Melanoma_Complexes %>% filter(dist < median(Melanoma_null_dist$value))

# Manually select Complexes to Plot and make plot 
write.csv(Melanoma_Complexes,'../dat/Output_tables/MelanomaComplex.csv')

Mel_comp_list <- c("Ribosome, cytoplasmic","Kinase maturation complex 1","DGCR8 multiprotein complex")

Mel_comp_plot <- Melanoma_null_dist

for (X in Mel_comp_list){
  
  comp_list <- unlist(str_split(comps$subunits.UniProt.IDs.[comps$ComplexName == X],';'))
  GO_mat <- t7_Melanomas %>% filter(rownames(t7_Melanomas) %in% comp_list)
  GO_mat <- as.matrix(GO_mat)
  GO_cor <- Distance_euclid_py(GO_mat,10)
  GO_cor$Condition <- X
  GO_cor <- GO_cor %>% dplyr::select(Condition,value)
  Mel_comp_plot <- rbind(Mel_comp_plot,GO_cor)
  
}


comp_mel <- ggplot(Mel_comp_plot,aes(x = log10(value),y = Condition)) +
  geom_density_ridges(aes(fill = Condition),stat='binline' ,bins = 35,rel_min_height = 0.01,scale = 2,alpha = .0,color=NA) +
  scale_fill_manual(values = c("#808080", "#F6BE00","#F6BE00","#F6BE00")) +
  theme(legend.position = "none")+ 
  geom_density_ridges(data=subset(Mel_comp_plot, Condition %in% Mel_comp_list[1]),
                      aes(fill = Condition),jittered_points = TRUE,point_alpha=.01,rel_min_height = 0.01,scale = .3,alpha = .5)+
  theme_bw() + coord_cartesian(xlim = c(-2,0)) + 
  ggtitle('Melanoma Complexes') + ylab('') + 
  xlab('Log10(Euclidean Distance)') +theme(legend.position = "none",axis.text=element_text(size=16),axis.title = element_text(size = 16),title = element_text(size = 18))+
  geom_density_ridges(data=subset(Mel_comp_plot, Condition %in% Mel_comp_list[2]),
                      aes(fill = Condition),jittered_points = TRUE,rel_min_height = 0.01,scale = .5,alpha = .5)+
  geom_density_ridges(data=subset(Mel_comp_plot, Condition %in% Mel_comp_list[3]),
                      aes(fill = Condition),jittered_points = TRUE,point_alpha=.5,rel_min_height = 0.01,scale = .3,alpha = .5)+
  geom_density_ridges(data=subset(Mel_comp_plot, Condition == ' All Proteins'),
                      aes(fill = Condition),jittered_points = FALSE,rel_min_height = 0.02,scale = .7,alpha = .7) + 
  coord_cartesian(xlim = c(-2,-.4))

comp_mel
```


Differential protein expression betweein Large and small subpopulations
```{r Diff protein expression}

pvals <- c()
meds <- c()
prot <- c()
for (i in 1:nrow(t7_Melanomas_small)){
  if(sum(is.na(t7_Melanomas_small[i,])==F) > 9){
    pvals <- c(pvals,wilcox.test(as.matrix(t7_Melanomas_small[i,]), as.matrix(t7_Melanomas_large[i,]), alternative = "two.sided")$p.value)
    prot <- c(prot,rownames(t7_Melanomas_small)[i])
    meds <- c(meds, median(as.matrix(t7_Melanomas_small[i,]),na.rm = T)- median(as.matrix(t7_Melanomas_large[i,]),na.rm = T))
    
  }
  
}




df <- as.data.frame(pvals)
df$prot <- prot
df$FC <- meds
df$qval <- p.adjust(df$pvals,method = 'BH')
df <- df %>% filter(qval < .01)
df$Condition <- NA
df$Condition[df$FC > 0] <- 'Cluster B'
df$Condition[df$FC < 0] <- 'Cluster A'

#write.csv(df,'../dat/Output_tables/diff_prot_subPOP.csv',row.names = T)

prots <- c('Q9Y261','P68363','P58546','Q99986','P14927','P53801','P49755')
genes <-c('FOXA2','TUBA1B','MTPN','VRK1','UQCRB','PTTG1IP','TMED10')

prots <- c('P68363','Q99986','P53801','P63241')
genes <-c('TUBA1B','VRK1','PTTG1IP','EIF5A')

t7_Melanomas_Small_plot <- t7_Melanomas_Small %>% filter(rownames(t7_Melanomas_Small)%in% prots)
t7_Melanomas_Small_plot <- t7_Melanomas_Small_plot[prots,]
t7_Melanomas_Small_plot$genes <- genes
t7_Melanomas_Small_plot <- melt(t7_Melanomas_Small_plot,id.vars = 'genes')
t7_Melanomas_Small_plot$Condition <- 'C2'

t7_Melanomas_Largel_plot <- t7_Melanomas_Large %>% filter(rownames(t7_Melanomas_Small)%in% prots)
t7_Melanomas_Largel_plot <- t7_Melanomas_Largel_plot[prots,]
t7_Melanomas_Largel_plot$genes <-genes
t7_Melanomas_Largel_plot <- melt(t7_Melanomas_Largel_plot,id.vars = 'genes')
t7_Melanomas_Largel_plot$Condition <- 'C1'


mel_dif_plot <- rbind(t7_Melanomas_Largel_plot,t7_Melanomas_Small_plot)
mel_dif_plot <- mel_dif_plot %>% filter(is.na(value)==F)


diff_prot_exp <- ggplot(data=mel_dif_plot, aes(x=value, y=genes)) +
  geom_density_ridges(aes(fill=Condition),alpha=.9, scale=1,stat = "binline", bins = 20) +
  theme_pubr() +
  scale_fill_manual(values=c('#DCDCDC','#808080')) +
  scale_x_continuous(limits=c(-3,3)) +
  coord_flip() +
  xlab("Protein level, log2") +
  ylab("Gene ontology") +
  rremove("y.ticks") +
  #rremove("x.text") +
  rremove("xlab") +
  font("ylab", size=30) +
  font("xy.text", size=30) +
  rremove("legend") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


mel_dif_plot_FOX_MTPN <- mel_dif_plot %>% filter(genes %in% c('MTPN','FOXA2'))
mel_dif_plot_FOX_MTPN <- dcast(mel_dif_plot_FOX_MTPN,Condition+variable~genes,value.var ='value' ,fill = NA)

p2 <- ggplot(mel_dif_plot_FOX_MTPN,aes(x = MTPN,y =FOXA2 ,color = Condition,shape = Condition,size = 15)) + 
  geom_point() + theme_classic() +scale_color_manual(values=c('#dcdcdc','#2C2C2C')) + rremove('legend')+ 
  coord_cartesian(xlim = c(-2,2))+theme(axis.title = element_text(size = 24), axis.text = element_text(size = 22))

mel_dif_plot_FOX_MTPN_ <- mel_dif_plot %>% filter(genes %in% c('UQCRB','TMED10'))
mel_dif_plot_FOX_MTPN_ <- dcast(mel_dif_plot_FOX_MTPN_,Condition+variable~genes,value.var ='value' ,fill = NA)


p1 <- ggplot(mel_dif_plot_FOX_MTPN_,aes(x = UQCRB,y =TMED10 ,color = Condition,shape = Condition,size = 15)) + 
  geom_point() + theme_classic() +scale_color_manual(values=c('#dcdcdc','#2C2C2C')) + rremove('legend') +
  coord_cartesian(xlim = c(-2,2))+theme(axis.title = element_text(size = 24), axis.text = element_text(size = 22))



p1
p2
ggsave('/Users/andrewleduc/Desktop/nPOP_Paper/figs/SubPop/joint_dist.png',plot = p1+p2,height = unit(6, "cm"),width = unit(12, "cm"))
ggsave('/Users/andrewleduc/Desktop/nPOP_Paper/figs/SubPop/one_dist.png',plot = diff_prot_exp,height = unit(6, "cm"),width = unit(10, "cm"))



```

Here is some code that computes significant correlations for the monocytes and melanoma cells, this is for figure 3.



